Parallel Information Set Generation for Kriegspiel

### 2012-07-15 sbac/paper.tex
spell check

%\documentclass[times, 10pt,twocolumn]{article}
%\documentclass[conference]{IEEEtran}
\documentclass[10pt, conference, compsocconf]{IEEEtran} 
%\usepackage{latex8}
%\usepackage{times}
\usepackage{url,hyperref}
\usepackage{subfigure}
\usepackage{tikz}
\usepackage{examplep}
\usepackage{moreverb}
\usepackage{algorithm}
\usepackage{boxedminipage}
\usepackage[noend]{algpseudocode}
\usepackage{listings}
\usepackage{amsmath,amsfonts,latexsym,graphicx}
\lstset{language=C,numbers=left,basicstyle=\footnotesize,numberstyle=\footnotesize}
\usetikzlibrary{arrows,decorations.pathmorphing,backgrounds,fit,shapes.geometric}
\usepgflibrary{shapes.geometric}

@@

\begin{abstract}
Information Set Generation is the identification of the set of paths in an
imperfect information game tree that are consistent with a player's
observations.  The ability to reason about the possible game history is
critical to the performance of game-playing agents.  ISG represent a class of
combinatorial search problems which is computationaly intensive but difficult
to efficiently parallelize.

@@

Figure~\ref{fig:figure1} shows the true state of the game after nine moves by
each player.  Figure~\ref{fig:figure2} shows White's perspective of the game at
that poin. At move 9, Black attempts to block a
potential threat along the diagonal by moving a pawn to g6, then attempts to
escape or capture a threat at f7.  When both of these attempts fail, Black can
infer that his king is being threatened by a protected bishop or queen at f7.
The attempt to move to e7 is Black's way of finding out whether the threatening
piece is a queen or bishop.  When this move fails, black knows it is White's
queen and ultimately retreats to d8.  Note that the failed moves by White on
turns 3 and 4 are a result of being blocked by the pawn at e2.

@@

%\begin{figure}
%%\begin{listing}
%\begin{lstlisting}
%void generateInformationSet(uint16_t* trueState, uint16_t* possState, bool whiteMove, 
%	uint16_t* possHistory, uint16_t** levels, int depth)
%{
%  if (!samePawnTries(trueState, possState, whiteMove)) return;  
%  if (!sameCheckStatus(trueState, possState, whiteMove)) return; 
%  if (depth == maxdepth) { // Then we have found a solution
%	nSolutions++; return; // Report solution
%  }
%  uint16_t newPossState[16]; 
%  int nMoves = 0;
%  generateAttemptableMoves(possState, whiteMove, levels[depth], nMoves,false);
%  checkForCheck(possState, whiteMove, levels[depth], nMoves); // prunes if not equal
%  assert (nMoves < NMOVES); // 
%
%  if (whitePerspective == whiteMove) { // move from perspective plyaer 
%    SetMove& failures = failedMoves[depth];
%    for (SetMove::const_iterator itr = failures.begin(); itr != failures.end(); ++itr) {
%      uint16_t move = *itr ;
%      if (!foundMatchingMove(move,levels[depth],nMoves)) {
%        return; // Prune: one of my own failed moves is not attemptable in this state
%      }
%    }
%    // We know exactly what the actual move was; make it
%    uint16_t& actualMove = moveHistory[depth];
%    if (!foundMatchingMove(actualMove,levels[depth],nMoves)) {
%      return; // Prune (because the move that we know we made at this depth is not legal)
%    }
%    // Otherwise, recurse 
%    applyMove(possState,newPossState,actualMove);
%    possHistory[depth] = actualMove;
%    generateInformationSet(newTrueState, newPossState, !whiteMove, 
%      possHistory, levels, depth+1);
%  } else { 
%    // We are at a level in the search tree where 
%    // we are considering the possible moves for the opponent.
%    // We know the number of attempted illegal moves, but not which ones
%
%    // Must be enough attemptable moves in the state to match the number of failed attempts
%    unsigned nIllegalMoves = countIllegalMoves(levels[depth], nMoves);
%    if (nIllegalMoves < failedMoves[depth].size()) {
%	return; // Prune if not enough attemptable moves
%    }
%    // Now we want to try each possible move that is legally executable (not just attemptable) 
%    for (int i = 0; i < nMoves; i++) {
%      uint16_t& move = levels[depth][i];
%      if (isLegal(move)) { // Obviously, we can only execute the moves that are actually legal from this state
%        applyMove(possState,newPossState,move);
%        possHistory[depth] = move;
%        generateInformationSet(newTrueState, newPossState, !whiteMove, 
%          possHistory, levels, depth+1);
%      }
%    }
%  }
%}
%\end{lstlisting}
%\caption{The heart of the information set generation algorithm.}
%\label{codelisting}
%\end{figure}

@@

\subsection{Analysis using Random Instance} 
The performance results reported in this section were obtained by running our
program on Blueprint with up to 1,024 processors. This system was composed of
120 POWER5+ nodes, each having 16 cores and 64GB of memory. The system was
decomissioned in June 2011 and is no longer in operation.

@@

\begin{table}[ht]
\centering
\begin{tabular}{rrr}
{\bf Procs}	&	{\bf Time (s)} 	&	{\bf Speedup}\\
1	&	1524	&	1\\
2	&	812	&	1.88\\
4	&	407	&	3.74\\
8	&	208	&	7.32\\
16	&	102	&	14.8\\
32	&	53.2	&	28.7\\
64	&	28.0	&	54.5\\
128	&	14.9	&	102\\
256	&	7.70	&	197\\
512	&	4.59	&	331\\
1024	&	2.64	&	576\\
\end{tabular}

\caption{Speedups for a randomly selected problem instance on 1--1,024
processors.  The input observations were from the 10th ply of the game tree and
the size of the information set in this instance is 4,495,121.}

@@

\begin{table}[ht]
\centering
\begin{tabular}{rrrrr}
PE/G & 2 & 4 & 6 & 8 \\
1 & {\bf 177$|$1} & 178$|$0.99 & 1844$|$0.96 & 493$|$0.36 \\
4 & 71$|$2.48 & 53$|$3.34 & {\bf 48$|$3.63} & 75$|$2.35 \\
16 & 34$|$5.19 & 19 $|$9.27& {\bf 12$|$13.8} & 16$|$10.7 \\
64 & 24 $|$7.25 & 7 $|$23.6& {\bf 3.7$|$47.39} & 4.1$|$43.1 \\
256 & 24 $|$ 7.1& 4.4 $|$40.1 & 1.2 $|$145& {\bf 1.0$|$172} \\
1024 & 26 $|$6.61 & 3.8 $|$46.5& 0.64 $|$276& {\bf 0.40$|$443}
\end{tabular}

\caption{Running times in seconds and speedups (after $|$), to find full
information set of size 764,209 at position 10 of Virgil-David 1969. Optimal
grainsize (G) for each number of cores (PE) is shown in bold.}

@@

\begin{table}[th]
\centering
\begin{tabular}{rrrrr}
PE/G & 2 & 4 & 6 & 8 \\
1 & 617$|$1 & {\bf 612$|$1} & 650$|$0.95 & 3312$|$0.19 \\
4 & 213$|$2.89 & 170$|$3.63 & {\bf 158$|$3.89} & 367$|$1.68 \\
16 & 83$|$7.41 & 49$|$12.4 & {\bf 40$|$15.1} & 64$|$9.57 \\
64 & 54$|$11.3 & 19$|$32.3 & {\bf 11$|$52.9} & 14.9$|$41.4 \\
256 & 54$|$11.3 & 9.8$|$62.9 & 3.5$|$172 & {\bf 3.5$|$174} \\
1024 & 58$|$10.6 & 5.5$|$111 & 1.3$|$464 & {\bf 1.2$|$536} \\
\end{tabular}

\caption{Running times, in seconds and speedups (after $|$), to find full
information set of size 5,099,257 at position 10 of Game 6 Kbot-Darkboard 2006.
Optimal grainsize for each number of cores is shown in bold.}
\label{tab:prob5}
\end{table}

We also evaluated our information set generation routine on 40 positions from
actual Kriespiel games, including human vs. human and computer vs. computer
instances.  Many of these were trivial to solve in a few seconds or less on a
desktop computer.  For these experiments, we used the Ranger supercomputer
at TACC~\cite{ranger}. Ranger has 16-way SMP nodes containing AMD Opteron
processors. The nodes are interconnected using Infiniband technology in a
fully-CLOS topology providing a 1GBps point-to-point bandwidth.

@@

\begin{table}[ht]
\centering
\begin{tabular}{lrrc}
{\bf Authors} & {\bf Game} & {\bf Year} & {\bf Best Speedup (cores)} \\
Ferguson \& Korf & Othello & 1988 & 12(32) \\
Felton \& Otto & Chess & 1988 & 100(256) \\
Richards \etal & Kriegspiel & 2012 & 178(256) \\
Richards \etal & Kriegspiel & 2012 & {\bf 536(1024)} \\
\end{tabular}
\caption{Best Speedups reported for game tree applications.}
\label{bestspeedups}
\end{table}

>>>>

%\documentclass[times, 10pt,twocolumn]{article}
%\documentclass[conference]{IEEEtran}
\documentclass[10pt, conference, compsocconf]{IEEEtran} 
%\usepackage{latex8}
%\usepackage{times}
\usepackage{url,hyperref}
\usepackage{subfigure}
\usepackage{tikz}
\usepackage{examplep}
\usepackage{moreverb}
\usepackage{algorithm}
\usepackage{boxedminipage}
\usepackage[noend]{algpseudocode}
\usepackage{listings}
\usepackage{amsmath,amsfonts,latexsym,graphicx}
\usepackage[tableposition=top,font=bf]{caption}

\lstset{language=C,numbers=left,basicstyle=\footnotesize,numberstyle=\footnotesize}
\usetikzlibrary{arrows,decorations.pathmorphing,backgrounds,fit,shapes.geometric}
\usepgflibrary{shapes.geometric}

@@

\begin{abstract}
Information Set Generation is the identification of the set of paths in an
imperfect information game tree that are consistent with a player's
observations.  The ability to reason about the possible game history is
critical to the performance of game-playing agents.  ISG represent a class of
combinatorial search problems which is computationally intensive but difficult
to efficiently parallelize.

@@

Figure~\ref{fig:figure1} shows the true state of the game after nine moves by
each player.  Figure~\ref{fig:figure2} shows White's perspective of the game at
that coin. At move 9, Black attempts to block a
potential threat along the diagonal by moving a pawn to g6, then attempts to
escape or capture a threat at f7.  When both of these attempts fail, Black can
infer that his king is being threatened by a protected bishop or queen at f7.
The attempt to move to e7 is Black's way of finding out whether the threatening
piece is a queen or bishop.  When this move fails, black knows it is White's
queen and ultimately retreats to d8.  Note that the failed moves by White on
turns 3 and 4 are a result of being blocked by the pawn at e2.

@@

%\begin{figure}
%%\begin{listing}
%\begin{lstlisting}
%void generateInformationSet(uint16_t* trueState, uint16_t* possState, bool whiteMove, 
%	uint16_t* possHistory, uint16_t** levels, int depth)
%{
%  if (!samePawnTries(trueState, possState, whiteMove)) return;  
%  if (!sameCheckStatus(trueState, possState, whiteMove)) return; 
%  if (depth == maxdepth) { // Then we have found a solution
%	nSolutions++; return; // Report solution
%  }
%  uint16_t newPossState[16]; 
%  int nMoves = 0;
%  generateAttemptableMoves(possState, whiteMove, levels[depth], nMoves,false);
%  checkForCheck(possState, whiteMove, levels[depth], nMoves); // prunes if not equal
%  assert (nMoves < NMOVES); // 
%
%  if (whitePerspective == whiteMove) { // move from perspective player 
%    SetMove& failures = failedMoves[depth];
%    for (SetMove::const_iterator itr = failures.begin(); itr != failures.end(); ++itr) {
%      uint16_t move = *itr ;
%      if (!foundMatchingMove(move,levels[depth],nMoves)) {
%        return; // Prune: one of my own failed moves is not attemptable in this state
%      }
%    }
%    // We know exactly what the actual move was; make it
%    uint16_t& actualMove = moveHistory[depth];
%    if (!foundMatchingMove(actualMove,levels[depth],nMoves)) {
%      return; // Prune (because the move that we know we made at this depth is not legal)
%    }
%    // Otherwise, recurse 
%    applyMove(possState,newPossState,actualMove);
%    possHistory[depth] = actualMove;
%    generateInformationSet(newTrueState, newPossState, !whiteMove, 
%      possHistory, levels, depth+1);
%  } else { 
%    // We are at a level in the search tree where 
%    // we are considering the possible moves for the opponent.
%    // We know the number of attempted illegal moves, but not which ones
%
%    // Must be enough attemptable moves in the state to match the number of failed attempts
%    unsigned nIllegalMoves = countIllegalMoves(levels[depth], nMoves);
%    if (nIllegalMoves < failedMoves[depth].size()) {
%	return; // Prune if not enough attemptable moves
%    }
%    // Now we want to try each possible move that is legally executable (not just attemptable) 
%    for (int i = 0; i < nMoves; i++) {
%      uint16_t& move = levels[depth][i];
%      if (isLegal(move)) { // Obviously, we can only execute the moves that are actually legal from this state
%        applyMove(possState,newPossState,move);
%        possHistory[depth] = move;
%        generateInformationSet(newTrueState, newPossState, !whiteMove, 
%          possHistory, levels, depth+1);
%      }
%    }
%  }
%}
%\end{lstlisting}
%\caption{The heart of the information set generation algorithm.}
%\label{codelisting}
%\end{figure}

@@

\subsection{Analysis using Random Instance} 
The performance results reported in this section were obtained by running our
program on Blueprint with up to 1,024 processors. This system was composed of
120 POWER5+ nodes, each having 16 cores and 64GB of memory. The system was
decommissioned in June 2011 and is no longer in operation.

@@

\begin{table}[ht]

\caption{Speedups for a randomly selected problem instance on 1--1,024
processors.  The input observations were from the 10th ply of the game tree and
the size of the information set in this instance is 4,495,121.}
\centering
\begin{tabular}{rrr}
{\bf Procs}	&	{\bf Time (s)} 	&	{\bf Speedup}\\
1	&	1524	&	1\\
2	&	812	&	1.88\\
4	&	407	&	3.74\\
8	&	208	&	7.32\\
16	&	102	&	14.8\\
32	&	53.2	&	28.7\\
64	&	28.0	&	54.5\\
128	&	14.9	&	102\\
256	&	7.70	&	197\\
512	&	4.59	&	331\\
1024	&	2.64	&	576\\
\end{tabular}

@@

\begin{table}[ht]
\caption{Running times in seconds and speedups (after $|$), to find full
information set of size 764,209 at position 10 of Virgil-David 1969. Optimal
grainsize (G) for each number of cores (PE) is shown in bold.}

\centering
\begin{tabular}{rrrrr}
PE/G & 2 & 4 & 6 & 8 \\
1 & {\bf 177$|$1} & 178$|$0.99 & 1844$|$0.96 & 493$|$0.36 \\
4 & 71$|$2.48 & 53$|$3.34 & {\bf 48$|$3.63} & 75$|$2.35 \\
16 & 34$|$5.19 & 19 $|$9.27& {\bf 12$|$13.8} & 16$|$10.7 \\
64 & 24 $|$7.25 & 7 $|$23.6& {\bf 3.7$|$47.39} & 4.1$|$43.1 \\
256 & 24 $|$ 7.1& 4.4 $|$40.1 & 1.2 $|$145& {\bf 1.0$|$172} \\
1024 & 26 $|$6.61 & 3.8 $|$46.5& 0.64 $|$276& {\bf 0.40$|$443}
\end{tabular}

@@

\begin{table}[th]

\caption{Running times, in seconds and speedups (after $|$), to find full
information set of size 5,099,257 at position 10 of Game 6 Kbot-Darkboard 2006.
Optimal grainsize for each number of cores is shown in bold.}
\centering
\begin{tabular}{rrrrr}
PE/G & 2 & 4 & 6 & 8 \\
1 & 617$|$1 & {\bf 612$|$1} & 650$|$0.95 & 3312$|$0.19 \\
4 & 213$|$2.89 & 170$|$3.63 & {\bf 158$|$3.89} & 367$|$1.68 \\
16 & 83$|$7.41 & 49$|$12.4 & {\bf 40$|$15.1} & 64$|$9.57 \\
64 & 54$|$11.3 & 19$|$32.3 & {\bf 11$|$52.9} & 14.9$|$41.4 \\
256 & 54$|$11.3 & 9.8$|$62.9 & 3.5$|$172 & {\bf 3.5$|$174} \\
1024 & 58$|$10.6 & 5.5$|$111 & 1.3$|$464 & {\bf 1.2$|$536} \\
\end{tabular}

\label{tab:prob5}
\end{table}

We also evaluated our information set generation routine on 40 positions from
actual Kriegspiel games, including human vs. human and computer vs. computer
instances.  Many of these were trivial to solve in a few seconds or less on a
desktop computer.  For these experiments, we used the Ranger supercomputer
at TACC~\cite{ranger}. Ranger has 16-way SMP nodes containing AMD Opteron
processors. The nodes are interconnected using Infiniband technology in a
fully-CLOS topology providing a 1GBps point-to-point bandwidth.

@@

\begin{table}[ht]
\caption{Best Speedups reported for game tree applications.}
\centering
\begin{tabular}{lrrc}
{\bf Authors} & {\bf Game} & {\bf Year} & {\bf Best Speedup (cores)} \\
Ferguson \& Korf & Othello & 1988 & 12(32) \\
Felton \& Otto & Chess & 1988 & 100(256) \\
Richards \etal & Kriegspiel & 2012 & 178(256) \\
Richards \etal & Kriegspiel & 2012 & {\bf 536(1024)} \\
\end{tabular}
\label{bestspeedups}
\end{table}
